---
title: "Figure arrangements"
author: "Pablo Gomez"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr); library(ggplot2); library(ggpubr);library(Pabloverse)
```

```{r import prev objects, include = F}
load('misc/II/captures.RData')
load('misc/II/figures.RData')
load('misc/II/tCaptures.RData')
```

# Objectives:
  
  - Talk about figure arrangements 
  - Make the 


# Arranging the plots in a layout

Now that we have all the figures in a list, we can make arrangements with our figures. For this we use the function `ggarrange()` from the ggpubr library.  

```{r}
library(ggpubr) # load the library
ggarrange(plotlist = figures)
```


# Facets

Facets are a way of stratifying the data based on variables in the data set, you can think about it in a similar way we have been using groups. To create a stratified plot we can use the function `facet_grid()` which will ask for a variable to go in the rows and another for columns:

```{r}
figures$histogram <- captures %>% # The data we will use 
  ggplot() + # set the canvas
  geom_histogram(aes(treated), fill = 'red4') + # We will create a histogram of the Age
  facet_grid(
    rows = vars(trap_type), # We will use the Sex variable for rows
    cols = vars(year) # We use the Result variable for Columns
  ) +
  theme_pubclean() # lets try another theme

figures$histogram
```


Now that you know some tools to look for information, you will have to make your figures on your own. If you want to make figures with the datasets we have been using you can do those, or you can use any of the code online to replicate them.  

1. Go to the [Data to Viz website](https://www.data-to-viz.com)
2. Identify a few figures that you would like to do (you can also do figures that we have previously done in the exercises)
3. Make sure you label your figures!
4. Experiment changing the colros, variables, themes etc...
5. Replicate a few figures and put them in an arrangement.

At the end we will have a discussion where you can share your figures and troubleshoot any problems you might have had.  


```{r Export objects, include=F}
# For now do it mannualy
save(captures, file = 'misc/III/captures.RData')
save(figures, file = 'misc/III/figures.RData')
save(tCaptures, file = 'misc/III/tCaptures.RData')
```

