---
title: "Topics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr); library(ggplot2); library(ggpubr);library(Pabloverse)
library(kableExtra)
```

# English

```{r echo=FALSE}
data.frame(
  Time = c('', '9:00-9:30', '9:30-10:30', '10:30-10:45', '10:45-12:00', '12:00-13:00', '13:00-14:00', '14:00-14:15', '14:15-15:00'),
  D01 = cell_spec(
    c('Topic', 'Introduction to the workshop', 'Data visualization with ggplot', 'Break', 'Further customization', 'Lunch', 'Figure arrangements', 'Break', 'Beyond static visualization'), 'html',
    link = c('', '', 'DataViz_I.html', 'coffeeBreak.html', 'DataViz_II.html', 'Lunch.html', 'DataViz_III.html', 'coffeeBreak.html', 'DataViz_IV.html')
  )
) %>%
  kbl("html", escape = FALSE, col.names = c('', 'June 10, 2023')) %>%
  kable_paper(full_width = T, lightable_options = c('hover')) %>%
  column_spec(1, bold = T, border_right = T) %>%
  row_spec(1, bold = T, extra_css = "border-bottom: 1px solid") %>%
  row_spec(c(4,6, 8), background = "#F6F5FA", italic = T)
```


# Spanish

```{r echo=FALSE}
data.frame(
  Time = c('', '9:00-9:30', '9:30-10:30', '10:30-10:45', '10:45-12:00', '12:00-13:00', '13:00-14:00', '14:00-14:15', '14:15-15:00'),
  D01 = cell_spec(
    c('Topic', 'Introduccion al curso', 'Visualizacion de datos con ggplot', 'Break', 'Mas opciones de customizacion', 'Lunch', 'Arreglos de figuras', 'Break', 'Mas allá de visualizaciones estáticas'), 'html',
    link = c('', '', 'ES/DataViz_I.html', 'coffeeBreak.html', 'ES/DataViz_II.html', 'Lunch.html', 'ES/DataViz_III.html', 'coffeeBreak.html', 'ES/DataViz_IV.html')
  )
) %>%
  kbl("html", escape = FALSE, col.names = c('', 'June 10, 2023')) %>%
  kable_paper(full_width = T, lightable_options = c('hover')) %>%
  column_spec(1, bold = T, border_right = T) %>%
  row_spec(1, bold = T, extra_css = "border-bottom: 1px solid") %>%
  row_spec(c(4,6, 8), background = "#F6F5FA", italic = T)
```

-------

# Sections:
  
  - *Introduction to the workshop*, Introduce instructors, participants, format and the schedule. Talk about the website and resources we will use. Maybe talk about some very generic stuff regarding data visualization without talking about ggplot. Take some time to also introduce some basic programming concepts (R and markdown 101)
  - *Data visualization with ggplot2*, Introduce the ggplot syntax and some basic types of figures:
    - Visualizing distributions, bar charts, histograms, box plots, other size based (i.e. bubble, pie)
    - Visualizing relationships, scatterplots, mozaic,
  - *Further customization* Color and shape, fills and other dimensions of visualization
  - *Figure arrangements* talk about different ways of making arrangements (Facets and ggarrange) and make the data viz exercise
  - *Beyond static visualization* Introduce to interactive stuff with plotly.