---
title: "Introduccion a la sintaxis de R"
author: "Pablo Gomez"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr); library(ggplot2); library(ggpubr);library(Pabloverse)
```

# Sintaxis en R

R es como una calculadora, podemos realizar diferentes operaciones matematicas. Por ejemplo:

```{r basic operations}
2 + 2
```

R es un lenguaje de programacion orientado a objetos. Esto significa que guardamos nuestra informacion en contenedores que llamamos objetos.  
En R podemos obtener el mismo resultado usando diferentes aproximaciones. Por ejemplo, para crear un objeto podemos escribir el nombre del objeto y asignar un valor usando los operadores `=` o `<-`. Podemos usar los objetos que creemos para realizar operaciones con otros objetos compatibles:

```{r storing values}
x = 2 # creamos un objeto con el operador = 
y <- 2 # creamos un objeto con el operador <- 
x + y # Hacemos operacions con los objetos
```

## Vectores en R

Podemos guardar mas de un valor en un objeto usando vectores. Los vectores en R son una secuencia de valores, para crear un nuevo vector podemos combinar múltiples valores con la función `c()`. 
Por ejemplo, vamos a crear una secuencia de números del 5 al 10 usando 2 aproximaciones diferentes, y después le preguntaremos a R si los objetos son iguales. 
**tip:** usando el atajo `alt + -` automáticamente agregara el operador `<-`. La decision sobre el operador para asignar objetos es cuestión personal, a mi en lo personal se me facilita usar el operador `<-` para asignar objetos ya que el operador `=` se usa también para otras cosas como veremos mas adelante, pero siéntete libre de usar el que se te haga mas intuitivo.

```{r vectors}
x <- c(5, 6, 7, 8, 9, 10) # Creamos una secuencia de números del 5 al 10
y = 5:10 # Creamos la misma secuencia, pero de manera diferente
x == y # Le preguntamos a R si son iguales
```

Observa que en el ejemplo anterior usamos el operador `==`, esta es una manera de decirle a R que queremos comparar, y no asignar, dos valores (o objetos). Recuerda que al usar un solo `=` asignaremos el valor de la derecha al de la izquierda.

### Indexing

Cuando tenemos un vector, podemos pedirle a R que nos de valores especificos en base a su posicion usando el operador `[ ]` y especificando dentro el numero deseado. Por ejemplo:

```{r vector operations}
x[3] # Aqui le pedimos el 3er valor del vector
x[3] * y[5] # Aqui multiplicamos el 3er valor por el 5to valor de nuestros vectores x y y
```

## Funciones

R has a lot of base functions, but we can define new functions. When using R studio, the key *Tab* will help us to auto complete, this can help us a lot when we don't remember the exact name of the functions available.
The best part of programming with R is that it has a very active community. Since its open source, anyone can create functions and compile them in a package (or library). we can download these packages and access new functions.  
Functions in R require arguments, which we can see in the function documentation or if we press the key *Tab* when we are inside the function.  

```{r functions}
# To get the sum of a vector of numbers inside an object we use sum()
sum(x)
```

We can put functions inside function, for example, to get $\sqrt{\sum_1^n x}$ the square root of a sum of the numbers in x we can use:

```{r nested functions}
sqrt(sum(x))
```

Making functions in R is not as complicated as it sounds and can be very useful when we need to do repetitive work. To define a function we need to include the arguments that we want for the function and what are we doing with those arguments. For example, the following function has only one argument which is a name (string) and just pastes some text before and after:

```{r creating a function}
F1 <- function(name){
  paste("Hola", name, "! welcome to the R world (: !") # paste the name with some text
}

# Testing the function (Put your name)
F1(name = "Pablo")
```

Besides storing numbers in the objects in R, we can store text, matrices, tables, spatial objects, images, and other types of objects. Next we will import our own data and do some manipulation in R.

**Exercise:** Create a function that performs the $\sqrt{\sum_1^n x}$ operation you did previously with the code `sqrt(sum(x))`

```{r Excercise 1, eval=FALSE, include=FALSE}
# Define the function
F2 <- function(x){
  sqrt(sum(x))
}

# Test the function
F2(x)
```

## Types of variables

There are different types of variables in R, so far we have used the numeric and string types. If yo want to know what kind of variable is a given object, you can use the function `class()`. Lets try it.

```{r variables I}
class(x)
```

**Exercise: ** What kind of variable is the output from the first function you defined `F1()`?

```{r exercise vars, eval=FALSE, include=FALSE}
class(F1(name = 'Pablo'))
```

The most commonly used types of variables include:

  - *numeric* This are continuous numeric variables WITH any decimal values. For example: KG of product imported, probability of an event happening.
  - *integer* These variables are whole numbers WITHOUT decimal values. For example: Number of animals, number of shipments, etc..
  - *character* Alphanumeric variables. For example: name of a region, name of a disease, farm ID.
  - *factor* Alphanumeric variable with specific categories or levels. For example: type of product imported, type of farm, etc...

## Introducing the pipes

The library `dplyr` has several functions that can help to clean, create new variables, and modify our data in other ways.  

```{r message = F}
# if we dont have installed the library we will need to install it using:
# install.packages("dplyr")
# we load the library:
library(dplyr)
```

`dplyr` introduces a new operator called *pipes* (`%>%`), which can connect several functions to an object. This is an alternative to write several functions in a single "line of code" in a more organized way. For example, if we want to execute a function `F1()` followed by another function `F2()` for the object `x`:  

`F2(F1(x))` is equivalent to `x %>% F1() %>% F2()`    

As you can notice, to read the code `F2(F1(x))` we have to go from the inside to the outside to see the order of execution of the functions, but when we read `x %>% F1() %>% F2()` we read from left to right, which is the same way we normally would read any text in western language.    

**Suggestion:** we can use the keys *Ctrl* + *shift* + *m* to insert the `%>%` operator.  

```{r pipes} 
# We previously used this code to calculate the square root of the sum of x
sqrt(sum(x))
```

Using the pipes we can do the same more organized, by writing the order of execution from left to right. 

```{r pipes II}
x %>% # First we call the data
  sum() %>% # Sum all the values
  sqrt() # Compute the square root
```

You will notice that the outputs are exactly the same. Feel free to use whatever syntax you prefer, but for this course we will use mostly the pipes and writing the code from left to right.

# Data manipulation

R can import data in different formats. The most common are excel files (*.csv*, *.xls* y *.xlsx*), text files *.txt* and spatial data *.shp*, which we will talk about more in detail later.  
To read *.xls*, *.xlsx* and *.shp* files we will need to install some libraries. To install a new library you need to be connected to the internet and use the function `install.packages()` to install the library. Once it has been installed, you can load the library using the function `library()`.  
  
Download the excel file from this [link](https://ucdavis.app.box.com/folder/174771408063?s=cltv7bcwblxw5v90z5l56ognt9dprehv). It's not necessary to have a Box account.  
  
*Suggestion:* Sometimes when we want to use only one function from a library, we can just write the name of the library followed by the operator `::` and the name of the function, for example: `package::function()`. This way we won't have to load the whole library.  

```{r}
# If we don't have the library installed, we use:
# install.packages("readxl")

# Import the excel file
PRRS <- readxl::read_xlsx("../Data/PRRS.xlsx")
```

The most popular format for tables in R are called `data.frame`, when we import the data from a *.csv* o *.xlsx*. We can examine what kind of object it is using the function `class()`:  

```{r}
# Ask the class of a given object
class(PRRS)
```

It is possible that objects can have more than 1 class. In the previous example, you can notice that the `PRRS` object has 3 different types of class.  

In the following section we will cover some basics of data manipulation, this includes create, modify and summarise variables in our data. 

## Reducing our data

Sometimes we want to select specific columns and rows on our data to reduce the dimensionality, for this we can use the functions:
  
  - `select()` to select specific *columns*
  - `slice()` to select specific *rows* based on position
  - `filter()` to select specific *rows* based on a condition
  
### Selecting specific columns

```{r select 1}
PRRS %>%  # the name of our data
  select(Result, farm_type) # we want to select only the columns Result and farm_type
```

We can also specify which columns we DON'T want to show in our data:

```{r select 2}
PRRS %>% 
  select(-Age, -id) # with a '-' before the name we will exclude the column from the data
```

### Filtering specific rows

Filtering columns works based on boolean logic, so we can specify a condition and R will show only the rows that satisfy that condition. For example, lets filter only the observations from boar studs: 

```{r filtering observations}
PRRS %>% 
  filter(farm_type == 'boar stud') # we will use the equals to operator for this
```

## Creating or modifying variables

To create a new variable we can use the function `mutate()`. For example, lets create a new variable that will tell us if the farm type is a sow farm or not. For this we use the variable `farm_type` which already contains information for different farm types 

```{r creating new variables}
# Creating a new variable
PRRS %>% # name of the data set
  mutate( # mutate is the function we sue to create new variables
    SowFarm = ifelse(farm_type == 'sow farm', 1, 0) # we will create a binary variable where 1 = sow farm, 0 = Any other farm type
  ) 
```

Notice that in the previous code chunk, we did not assigned it to an object, which means that R is just going to print it and dont save it. To save it, remember that we use the assign operator `<-` or `=`.  
To modify an existing variable we can use the same function `mutate()`, and specify what we want to replace the existing variable with. For example, in the next chunk of code, we will modify the variable *Sex*. The original variable has H for Female and M for Male, so let's change it to something that makes more sense: 

```{r modifying an existing variable}
# Now we will assign the new object to itself overwriting it.
PRRS <- PRRS %>%  # this is the data set we will use
  mutate( # we use the mutate function to create new variables
    Sex = recode( # The function recode, can be used similar to replace in excel
      Sex, # we will overwrite the variable Sex using the same name
      H = 'Female',
      M = 'Male'
    )
  )
```

Be careful when overwriting objects in R there is no undo for this. It is important that your code is ordered and replicable in case you make any mistake, so you can easily reach to your progress before the mistake.  

## Grouping and summarizing tables

Often times we want to calculate summary statistics from our data, we can group by a specific variable (or even multiple variables) to examine if there are differences between groups.  

The most basic way of doing this is with the function `count()` which will only tell us the number of rows for each group:

```{r count}
# We can count how many observations for each of a variable name
PRRS %>% 
  count(Result)
```

The previous table show us that were 986 observations with test result negative and 367 with test result positive. We can add multiple variables to count. Now we will count by *Result* and *Sex*:

```{r count2}
# We can count based on multiple variables
PRRS %>% 
  count(Result, Sex)
```

We can also calculate other statistics by group. For example lets calculate the mean and standard deviation of the age by Result and Sex:

```{r}
PRRS %>%
  group_by(Result, Sex) %>% # The groups used for the data
  summarise( # the function summarise calculates statistics by the defined groups
    meanAge = mean(Age), # Calculate the mean age
    sdAge = sd(Age) # Calculate the standard deviation
  )
```

```{r weighted count}
# PRRS %>% 
  # count(Sex, wt = Result)
```

## Joining data sets

Sometimes we have different data sets that have variables in common and we want to integrate them into a single data set for further analysis. In this example we have the data sets *node_attrib.csv* and *network.csv*

```{r}
# First we import the data -----------
# Importing the farm dataset 
nodes <- read.csv("../Data/node_attrib.csv")
# Importing the movement dataset
mov <- read.csv("../Data/network.csv")
```

The data `mov` has information of place of origin (*id_orig*) and destination (*id_dest*) of animal shipments. First we will calculate the total number of pigs moved for each of the incoming and outgoing movements. To do this we will use the function `count()`

```{r}
# Get the number of outgoing and incoming shipments 
Out <- mov %>% # First we call the data
  count(id_orig, wt = pigs.moved) %>% # then we count the number of movements
  rename(id = id_orig, outgoing = n) # Rename the variables
```

You will notice that we added the argument `wt = pigs.moved`, *pigs.moved* is a variable that tell us the number of pigs shipped. The argument `wt` in the function `count()` allow us to add a weight to the counts. We also used the function `rename()` which pretty much does what it sounds like, it renames variables in the data.

Now we will join the data with the farm information contained in the *nodes* object, to do this we use the `left_join()` function:

```{r}
# Joining the two tables
# First we join with the outgoing
farms <- nodes %>% # This is the main object we will join with
  left_join(Out, by ="id") # we need to specify the variable name we are joining by
```

**Exercise: ** Now create a object that is the same as the `Out` we just created, but instead of counting the outgoing animals, counts the incoming animals. For this you can use the column *id_dest*. Then join the new object you created with the farms data. Name the variable that has the number of pigs moved *incoming*

```{r}
# Get the number of outgoing and incoming shipments 
In <- mov %>% # First we call the data
  count(id_dest, wt = pigs.moved) %>% # then we count the number of movements
  rename(id = id_dest, incoming = n) # Rename the variable

## Then we join with the incoming
farms <- farms %>%
  left_join(In, by = "id")
```

The first rows of your data should look like these:

```{r}
head(farms)
```

Notice that we have some outgoing and incoming rows with NA values. Sometimes NAs can interfer with some functions such as sum, in the next example we will summarise the data and egt the ones with the highest number of incoming movements:

```{r}
farms %>% # This is our joined data
  group_by(id) %>% # We will group it by id
  summarise( # we will perform some summary statistics
    incoming = sum(incoming, na.rm = T) # notice that we add the argument na.rm = T
  ) %>%
  arrange(desc(incoming)) %>% # now we will order by incoming
  head(5) # we use the head() function to get the first 5 observations
```

<!-- ## Wide and Long formats -->

<!-- # Loops in R -->

# Exporting the data

Once that you have processed the data, often times you want to export it so the next time dont have to run all the code. YOu can export to multiple formats, but the most common is to a comma delimited file (CSV), which can be read in excel.

```{r eval = F}
write.csv(farms, 'Data/farms.csv', row.names = F)
```


------------------------------------------------------------------------

*This lab has been developed with contributions from: [Jose Pablo
Gomez-Vazquez](https://www.spablo-temporal.network).\
Feel free to use these training materials for your own research and
teaching. When using the materials we would appreciate using the proper
credits. If you would be interested in a training session, please
contact: [jpgo\@ucdavis.edu](mailto:jpgo@ucdavis.edu){.email}*

------------------------------------------------------------------------